---
# This doesn't get used for the title slide...keep scroling!
title: "Presentation"
output:
  xaringan::moon_reader:
    seal: false
    df_print: paged
    lib_dir: libs
    css:
    - default
    - assets/theme.css
    - assets/theme-xaringan.css
    - presentation.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["theme-xaringan-title-slide", "center", "middle"]
---
class: theme-xaringan-title-slide, center, middle

```{r setup, include=FALSE}
library(sf)
library(s2)
library(s2plot)

if (!dir.exists("ne_countries")) {
  curl::curl_download(
    "https://naturalearth.s3.amazonaws.com/110m_cultural/ne_110m_admin_0_countries.zip",
    "ne_110m_admin_0_countries.zip"
  )
  dir.create("ne_countries")
  unzip("ne_110m_admin_0_countries.zip", exdir = "ne_countries")
  unlink("ne_110m_admin_0_countries.zip")
}

countries_s2 <- s2_data_countries()
countries_sf <- read_sf("ne_countries/ne_110m_admin_0_countries.shp") %>% 
  st_geometry() %>% 
  st_set_crs(NA)



plot_flat <- function(x, ...) {
  default_par <- list(mai = rep(0.5, 4), fg = "white", bg = "#26374a", col.axis = "white")
  withr::with_par(default_par, {
    wk::wk_plot(x)
    force(list(...))
  })
  invisible(NULL)
}

plot_round <- function(x, ..., lat = 0) {
  # hack to get preview and knit versions working
  if (!isTRUE(getOption("rstudio.notebook.executing"))) {
    bg <- NA
    frame_subset <- -1
  } else {
    bg <- "#26374a"
    frame_subset <- 1
  }
  
  s2plot_default_par <- list(mai = rep(0, 4), fg = "white", bg = bg, col.axis = "white")
  exprs <- rlang::enquos(...)
  
  withr::with_par(s2plot_default_par, {
    for (lon in seq(0, -360, length.out = 201)[frame_subset]) {
      pj <- s2plot_projection_orthographic(sprintf("POINT (%s %s)", lon, lat))
      pj_env <- s2plot:::last_projection_env
      pj_env$last_projection <- pj
      
      plot(double(), double(), xlim = c(-1, 1), ylim = c(-1, 1), asp = 1, axes = FALSE)
      s2plot(
        x,
        col = NA, border = "white", 
        projection = pj,
        add = TRUE
      )
      
      lapply(exprs, rlang::eval_tidy)

      wk::wk_plot(wk::crc(0, 0, r = 1), border = "grey70", add = T)
    }
  })
  invisible(NULL)
}

knitr::opts_chunk$set(
  dpi = 200,
  fig.height = 4,
  echo = FALSE
)

# this just has to be here for xaringan to work
options(htmltools.dir.version = FALSE)
```

<!-- this is actually the title slide --->

# Open source geometry on the sphere using s2geometry

### Dewey Dunnington<sup>1</sup> and Edzer Pebesma<sup>2</sup>

<br />
<br />
<br />

<sup>1</sup>Fisheries and Oceans Canada<br />
<sup>2</sup>Institute for Geoinformatics, University of Münster

<!-- this is how you get the parliament hill blue thing on your slide  --->
.landscape[

]

---
class: inverse, center, middle

# Your data is this:

### longitude: -63.56 ..., latitude: 45.11 ...

---
class: inverse, center, middle

# ...but the world is not this:

```{r countries-flat}
plot_flat(countries_sf)
```

---
class: inverse, center, middle

# it's this:

```{r countries-round, animation.hook='gifski', interval = 1 / 25}
plot_round(countries_s2)
```

---
class: inverse, center, middle

# Your question is this:

### Which {something} is within 500 km of land?

---
class: inverse, center, middle

# ...but the answer is not this:

```{r buffer-flat}
# 500 km is ~5 degrees at the equator
countries_sf_buffer <- countries_sf %>%
  st_combine() %>% 
  st_union() %>%
  st_buffer(5)


plot_flat(countries_sf_buffer)
```

---
class: inverse, center, middle

# it's this:

```{r buffer-round, animation.hook='gifski', interval = 1 / 25}
countries_s2_buffer <- countries_s2 %>% 
  s2_coverage_union_agg(s2_options(snap = s2_snap_level(15))) %>% 
  s2_buffer_cells(500000, min_level = 8) %>% 
  s2_rebuild(s2_options(snap = s2_snap_level(15)))

plot_round(countries_s2_buffer)
```

---
class: inverse, center, middle

# Your bounding box is not this:

```{r bbox-flat}
plot_flat(countries_sf, wk::wk_plot(wk::rct(-180, 51, 180, 72), add = TRUE))
```

---
class: inverse, center, middle

# it's this:

```{r bbox-round, animation.hook='gifski', interval = 1 / 25}
alaska_bbox_seg <- wk::rct(-180, 51, 180, 72) %>% 
  st_as_sfc() %>% 
  st_segmentize(1)

alaska_s2 <- s2_data_countries("United States of America") %>% 
  s2_intersection(alaska_bbox_seg)

alaska_bbox_s2 <- alaska_s2 %>% 
  s2_buffer_cells(0, max_cells = 4) %>%
  wk::as_wkb() %>%
  wk::wk_handle(
    s2::s2_unprojection_filter(
      s2::s2_projection_filter(
        tessellate_tol = 0.0001,
        wk::wkb_writer()
      )
    )
  ) %>% 
  as_s2_geography()

plot_round(
  countries_s2,
  s2plot(alaska_bbox_s2, col = rgb(1, 1, 1, 0.5), border = NA, add = T),
  lat = 75
)
```

---
class: inverse, center, middle

# or this:

```{r bbox-round-cap, animation.hook='gifski', interval = 1 / 25}
alaska_cap <- alaska_s2 %>% 
  s2_bounds_cap()
alaska_cap_geog <- s2_geog_point(alaska_cap$lng, alaska_cap$lat) %>% 
  s2_buffer_cells(alaska_cap$angle * pi / 180, radius = 1) %>% 
  s2_simplify(0.001, radius = 1)

plot_round(
  countries_s2,
  s2plot(alaska_cap_geog, col = rgb(1, 1, 1, 0.5), border = NA, add = T),
  lat = 75
)
```

---

# s2geometry

- s2geometry (<https://s2geometry.io>) is an open-source spherical geometry engine and indexing system created and supported by Google

- The R package **s2** provides flexible R bindings for s2geometry using an interface similar to PostGIS and BigQuery Geography

- As of **sf** version 1.0.0, s2geometry is the default engine for geodetic coordinates

---

# Using s2 in R via s2

```{r, echo = TRUE}
library(s2) # install.packages("s2")

cities <- as_s2_geography(s2_data_tbl_cities$geometry)
city_names <- s2_data_tbl_cities$name
canada <- s2_data_countries("Canada")

city_names[s2_intersects(cities, canada)]
```

---

# Using s2 in R via sf

```{r, echo = TRUE}
# good example?
```

---

# Challenges

```{r, echo = TRUE, error = TRUE}
nc <- read_sf(system.file("gpkg/nc.gpkg", package = "sf"))
nc %>% st_combine() %>% st_union()
```

---

# Acknowledgements

- s2geometry authors: Eric Veach, Jesse Rosenstock

- Fisheries and Oceans Canada: Clark Richards, Blair Greenan

- Edzer Pebesma for roping me into all of this!

---

# Questions?

.landscape[

]
